Document Expansion Using External Collections

### 2017-04-13 expansion/paper.tex
addressing some comments

\documentclass{sig-alternate}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{verbatim}
\usepackage{colortbl}
\usepackage[table]{xcolor}

@@

%\author{Garrick Sherman}
\author{[Blind]}

@@

\maketitle
\begin{abstract}
Document expansion has been shown to improve the effectiveness of information retrieval systems by augmenting documents' term probability estimates with those of similar documents, producing higher quality document representations. We propose a method to further improve document models by utilizing external collections as part of the document expansion process. Our approach is based on relevance modeling, a popular form of pseudo-relevance feedback; however, where relevance modeling is concerned with \textit{query} expansion, we are concerned with \textit{document} expansion. Our experiments demonstrate that the proposed model improves ad-hoc document retrieval effectiveness on a variety of corpus types, with a particular benefit on more heterogeneous collections of documents.
\end{abstract}

@@

%We can use this idea to expand the query representation into an estimate of the relevance model (see Section \ref{section.expanding.model.rm} for more details) and rank documents by the similarity of their estimated language model to the estimated RM.

@@

Document expansion has been well studied in information retrieval literature, e.g. \cite{Liu2004, Singhal1999, Tao2006, Wei2006}. For example, Liu \& Croft propose a method of retrieval that uses document clusters to smooth document language models \cite{Liu2004}. Tao et al. propose a similar approach but place each document at the center of its own cluster; this helps to ensure that the expansion documents are as closely related to the target document as possible \cite{Tao2006}.

@@

The incorporation of external collections into document retrieval is a similarly common theme in the ad-hoc IR literature, particularly with respect to query expansion \cite{Bendersky2012, Diaz2006, Li2007, Weerkamp2009, Xu2009}. Of particular relevance to our work is that of Diaz \& Metzler, whose mixture of relevance models is the basis of our Eq. \ref{eq.ql-and-expansion-mult} \cite{Diaz2006}. Their model simply interpolates RMs built on different collections, weighting each by a query-independent quantity $P(c)$. Though our work bears similarities, Diaz \& Metzler are interested in query expansion, whereas we apply the technique as one piece in a document expansion model.

@@

\noindent where $c(w, Q)$ is the frequency of word $w$ in $Q$. We follow standard procedures for estimating $P(w|\theta_D)$ in Eq. \ref{equation.ql}, estimating a smoothed language model by assuming that document language models in a given collection have a Dirichlet prior distribution:

@@

\noindent where $P(Q|D)$ is calculated as in Eq. \ref{equation.ql} and essentially weights word $w$ in $D$ by the query likelihood of the document. Relevance models are most efficient and robust when calculated over only the top ranked documents and limited to the top terms. These parameters are referred to as $fbDocs$ and $fbTerms$ respectively in Table \ref{table.parameters} below.

@@

To expand a document $D$, we begin by treating the text of $D$ as a pseudo-query which we pose against a collection of documents $C_E$.  To transform a document into a pseudo-query we apply two transformations.  First we remove all terms from $D$ that appear in the standard Indri stoplist\footnote{http://www.lemurproject.org/stopwords/stoplist.dft}.  Next, we prune our pseudo-query by retaining only the $0 < k \leq |D|$ most frequent words in the stopped text of $D$.  The integer variable $k$ is a parameter that we choose empirically.  Let $Q_D$ be the pseudo-query for $D$, consisting of the text of $D$ after our two transformations are applied.

We rank related documents, called expansion documents, by running $Q_D$ over an index $C_E$. More formally, we rank the documents in $C_E$ against $D$ using Eq. \ref{equation.ql}, substituting $Q_D$ for the query and $E_i$---the $i^{th}$ expansion document---for the document. Let $\pi_i$ be the log-probability for expansion document $E_i$ with respect to $D$ given by Eq. \ref{equation.ql}.

@@

Given our motivating intuition that document expansion allows for the more accurate estimation of document language models, we would expect that an RM computed using expanded documents should be more accurate than a standard RM. We therefore compute an RM3 as in Eqs. \ref{equation.rm1} and \ref{equation.rm3}, substituting the expanded document for the original.

@@

Although Eq. \ref{eq.ql-and-expansion-mult} allows for an arbitrary number of collections, for now we limit ourselves to two: the collection that the document appears in (the ``target'' collection) and Wikipedia\footnote{http://en.wikipedia.org}. We expect the latter, as a general encyclopedia, to yield relatively unbiased probability estimates. We build an Indri\footnote{http://www.lemurproject.org/indri/} index over the Wikipedia page text.

@@

Stop words are removed from the query. For efficiency, we retrieve the top 1000 documents using the default Indri QL implementation and re-rank these documents based on their expanded representations as described in Section \ref{section.model}.

@@

The parameters required for our approach, their meanings, and the values used in our experiments are shown in Table \ref{table.parameters}.

For this work, we set $k$ heuristically. In principle, this parameter need not be limited beyond the length of the document; however, this would increase computation time significantly, so we have set it to a reasonable value. The parameter $n$ is also set heuristically; see Section \ref{section.n-sensitivity} for a discussion of the sensitivity of our model to the setting of $n$.

@@

Retrieval effectiveness of each run is shown in Table \ref{table.effectiveness}. We measure effectiveness with mean average precision (MAP) and normalized discounted cumulative gain at 20 (nDCG@20). Each metric is optimized with 10-fold cross-validation.

@@

Figure \ref{figure.n-sweeps} shows sweeps over several values of $n$, the number of expansion documents, for the \textit{self} and \textit{wiki} QL runs using our established cross-validation procedure with identical folds. The sensitivity to $n$ is not pronounced at $n \geq 5$, and what little variation exists is not consistent across collections. We therefore set $n$ to 10, an apparently safe value, for all other runs. This is a convenient result since it allows for more efficient document expansion.

>>>>

\documentclass[sigconf]{acmart}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{verbatim}
\usepackage{colortbl}
%\usepackage[table]{xcolor}

@@

\author{Garrick Sherman}

\copyrightyear{2017}
\acmYear{2017}
\setcopyright{acmcopyright}
\acmConference{SIGIR '17}{}{August 7--11, 2017, Shinjuki, Tokyo, Japan}
\acmPrice{15.00.}
\acmDOI{http://dx.doi.org/10.1145/XXXXXXX.XXXXXXX}
\acmISBN{ISBN 978-1-4503-5022-8/17/08}

@@

\begin{abstract}
Document expansion has been shown to improve the effectiveness of information retrieval systems by augmenting documents' term probability estimates with those of similar documents, producing higher quality document representations. We propose a method to further improve document models by utilizing external collections as part of the document expansion process. Our approach is based on relevance modeling, a popular form of pseudo-relevance feedback; however, where relevance modeling is concerned with \textit{query} expansion, we are concerned with \textit{document} expansion. Our experiments demonstrate that the proposed model improves ad-hoc document retrieval effectiveness on a variety of corpus types, with a particular benefit on more heterogeneous collections of documents.
\end{abstract}

\maketitle

@@

%We can use this idea to expand the query representation into an estimate of the relevance model (see Section~\ref{section.expanding.model.rm} for more details) and rank documents by the similarity of their estimated language model to the estimated RM.

@@

Document expansion has been well studied in information retrieval literature, e.g.\ \cite{Liu2004, Singhal1999, Tao2006, Wei2006}. For example, Liu \& Croft propose a method of retrieval that uses document clusters to smooth document language models \cite{Liu2004}. Tao et al.\ propose a similar approach but place each document at the center of its own cluster; this helps to ensure that the expansion documents are as closely related to the target document as possible \cite{Tao2006}.

@@

The incorporation of external collections into document retrieval is a similarly common theme in the ad-hoc IR literature, particularly with respect to query expansion \cite{Bendersky2012, Diaz2006, Li2007, Weerkamp2009, Xu2009}. Of particular relevance to our work is that of Diaz \& Metzler, whose mixture of relevance models is the basis of our Eq.~\ref{eq.ql-and-expansion-mult} \cite{Diaz2006}. Their model simply interpolates RMs built on different collections, weighting each by a query-independent quantity $P(c)$. Though our work bears similarities, Diaz \& Metzler are interested in query expansion, whereas we apply the technique as one piece in a document expansion model.

@@

\noindent where $c(w, Q)$ is the frequency of word $w$ in $Q$. We follow standard procedures for estimating $P(w|\theta_D)$ in Eq.~\ref{equation.ql}, estimating a smoothed language model by assuming that document language models in a given collection have a Dirichlet prior distribution:

@@

\noindent where $P(Q|D)$ is calculated as in Eq.~\ref{equation.ql} and essentially weights word $w$ in $D$ by the query likelihood of the document. Relevance models are most efficient and robust when calculated over only the top ranked documents and limited to the top terms. These parameters are referred to as $fbDocs$ and $fbTerms$ respectively in Table~\ref{table.parameters} below.

@@

To expand a document $D$, we begin by treating the text of $D$ as a pseudo-query which we pose against a collection of documents $C_E$. To transform a document into a pseudo-query we apply two transformations. First we remove all terms from $D$ that appear in the standard Indri stoplist\footnote{http://www.lemurproject.org/stopwords/stoplist.dft}. Next, we prune our pseudo-query by retaining only the $0 < k \leq K$ most frequent words in the stopped text of $D$, where $K$ is the total number of unique terms in $D$. The integer variable $k$ is a parameter that we choose empirically. These are the non-stop words with the highest probabilities in a maximum likelihood estimate of $D$'s language model and are therefore a reasonable representation of the topic of the document. Though some information may be lost with stopping, with a large enough $k$ we hope to nevertheless capture the general topic of a document; for example, a document about Hamlet's famous speech may not be represented by the terms ``to be or not to be,'' but the terms ``Shakespeare,'' ``Hamlet,'' ``speech,'' etc. will likely represent the document's subject sufficiently. Let $Q_D$ be the pseudo-query for $D$, consisting of the text of $D$ after our two transformations are applied.

We rank related documents, called expansion documents, by running $Q_D$ over a collection $C_E$. More formally, we rank the documents in $C_E$ against $D$ using Eq.~\ref{equation.ql}, substituting $Q_D$ for the query and $E_i$---the $i^{th}$ expansion document---for the document. Let $\pi_i$ be the log-probability for expansion document $E_i$ with respect to $D$ given by Eq.~\ref{equation.ql}.

@@

Given our motivating intuition that document expansion allows for the more accurate estimation of document language models, we would expect that an RM computed using expanded documents should be more accurate than a standard RM. We therefore compute an RM3 as in Eqs.~\ref{equation.rm1} and~\ref{equation.rm3}, substituting the expanded document for the original.

@@

Although Eq.~\ref{eq.ql-and-expansion-mult} allows for an arbitrary number of collections, for now we limit ourselves to two: the collection that the document appears in (the ``target'' collection) and Wikipedia\footnote{http://en.wikipedia.org}. We expect the latter, as a general encyclopedia, to yield relatively unbiased probability estimates. We build an Indri\footnote{http://www.lemurproject.org/indri/} index over the Wikipedia page text.

@@

Stop words are removed from the query. For efficiency, we retrieve the top 1000 documents using the default Indri QL implementation and re-rank these documents based on their expanded representations as described in Section~\ref{section.model}.

@@

The parameters required for our approach, their meanings, and the values used in our experiments are shown in Table~\ref{table.parameters}.

For this work, we set $k$ heuristically. In principle, $k$ may equal the length of the document; however, this would increase computation time significantly, so we have set it to a smaller value for efficiency. The parameter $n$ is also set heuristically; see Section~\ref{section.n-sensitivity} for a discussion of the sensitivity of our model to the setting of $n$.

@@

Retrieval effectiveness of each run is shown in Table~\ref{table.effectiveness}. We measure effectiveness with mean average precision (MAP) and normalized discounted cumulative gain at 20 (nDCG@20). Each metric is optimized with 10-fold cross-validation.

@@

Figure~\ref{figure.n-sweeps} shows sweeps over several values of $n$, the number of expansion documents, for the \textit{self} and \textit{wiki} QL runs using our established cross-validation procedure with identical folds. The sensitivity to $n$ is not pronounced at $n \geq 5$, and what little variation exists is not consistent across collections. We therefore set $n$ to 10, an apparently safe value, for all other runs. This is a convenient result since it allows for more efficient document expansion.
